Possible Exoskeleton Control Architectures and Algorithms\\ (Draft 4.0)

### 2015-09-30 control.tex

We recommend a red team that explores frequency domain control system
design, as well as an emphasis on ``modern'' control in this paper.

@@

%\section{Symbiosis and Autonomy, Handling Errors, and Superhuman Reflexes}
A later white paper could discuss
possible combinations of response to user and exoskeleton
autonomy including balance control; response to trips, slips,
stumbles, and fumbles; task assistance and guidance (guide operator to
doorknob, button, or light switch); superhuman response to external
perturbations such as projectiles and explosions; and autonomous
execution (operator/exoskeleton symbiosis with multi-tasking)?

A later white paper could discuss how to handle exoskeleton
malfunctions and damage to the
system. Basically this boils down to fault detection, and switch over to
an appropriate safe mode of operation.

A later white paper could evaluate using simulation
the various control approaches' robustness to modeling error and unmodeled
dynamics.

@@

If you ask people who research complex computationally expensive
high end control
systems for a living how to control a system,
they will recommend complex computationally expensive high end control systems.
It is entirely possible that simple local low dimensional controllers
designed using traditional (frequency domain) techniques will do
quite well, especially if the appropriate sensors (such as 
operator-exoskeleton force sensors) are available.
It may even be the case that frequency domain techniques will handle
structural modeling error (jello-like human operators, deformable materials
(straps and pads, wires and hoses), and structural deformation and joint
play in lightweight frames) better than the state space oriented techniques
advocated here.

We recommend a red team that explores frequency domain control system
design, as well as an emphasis on ``modern'' control here.

@@

Typically low level control is implemented
by a high sampling rate global servo~\cite{Atlas-robot}
or distributed limb or joint-level high sampling rate servos~\cite{Sarcos-robot}.
Often, this controller is provided by the manufacturer of the robot
or exoskeleton and runs on a separate computer system from the
``user's'' or task/application controller~\cite{Sarcos,Atlas}.

@@

In this architecture, state estimation to deal with sensor noise,
disturbance estimation, and incomplete or
redundant sensing is a separate process and is as independent as possible
from the control method~\cite{certainty-eq,separation-prin}.
This makes evaluating different control methods much easier.
Disturbance observers for disturbances such as
moving support surfaces, contact or wind forces, and model errors
are also implemented.

@@

The recommended control architecture is intuitive, abstracts the hardware 
and lower level control in a straightforward way, is easy to program, and is
commonly used. Variants of this architecture were used by almost all teams
in the DARPA Robotics Challenge~\cite{}.

@@

An alternative approach to state estimation is to design state estimation
and control laws simultaneously~\cite{LQG-LTR}. Another alternative is
to have a controller with internal state, and design or learn a policy
that works well, but does not have an explicit state estimation 
part~\cite{lead-lag}.

@@

A key question is where does the desired acceleration $\ddot{\vq}_d$ come from?
This is discussed in a companion paper [intent.pdf].

@@

Full (operator) body
{\bf online inverse dynamics} as described in Section~\ref{sec:invdyn}
no longer works if $\mJ_{\myx}$ is not invertible (not square or not full rank).
In the underactuated and overactuated cases $\mJ_{\myx}$ is not square. 
For a partial (operator) body
exoskeleton we can rewrite the dynamics in terms of only those
joints the exoskeleton actuates, and invert those dynamics to cancel exoskeleton
effects on those joints. This is what has been done with the BLEEX exoskeleton,
for example.~\cite{}.

@@

Another approach to solving this problem is prioritized control, which does
not include optimization~\cite{}. For reasons we will describe, we believe
online optimization is much more flexible and can handle programming exoskeleton
behavior much better, so we will not discuss prioritized control further.

@@

Quadratic programming~\cite{Wikipedia} solves the following problem:
Minimize
\begin{equation}
\vx^{\tr} \mQ \vx + c^{\tr} \vx
\end{equation}
with inequality constraints
\begin{equation}
\mA_1 \vx \le \vb_1
\end{equation}
and equality constraints
\begin{equation}
\mA_2 \vx = \vb_2
\end{equation}
The optimization variable $\vx$ includes accelerations $\ddot{\vq}$, actuator
commands $\vu$, and contact forces $\vf$.
We could optimize only taking into account the exoskeleton and its dynamics,
minimizing the discrepancy between actual and desired accelerations, actuator command
magnitude, and contact force magnitude:
\begin{equation}
C_{\myx}(\vx) = (\vq - \vq_d)^{\tr} \mW_{\vq} (\vq - \vq_d)
+ \vu_{\myx}^{\tr} \mW_{\vu_{\myx}} \vu_{\myx}
+ \vf_{\ox}^{\tr} \mW_{\vf_{\ox}} \vf_{\ox}
+ \vf_{\xw}^{\tr} \mW_{\vf_{\xw}} \vf_{\xw}
\end{equation}
$\mW_{\vq}$, $\mW_{\vu_{\myx}}$, $\mW_{\vf_{\ox}}$,
and $\mW_{\vf_{\xw}}$ are weight matrices to trade off the
different errors/magnitudes and error/magnitude directions.
The dynamics (equation~\ref{eq:xdynamics2}) are enforced as an equality constraint.
Actuator limits, friction limits, center of pressure location constraints,
and unidirectional contact constraints are all
enforced as inequality constraints.
Variants of this optimization
were performed every 1-2ms by various teams to control the
Atlas robots in the DARPA Robotics Challenge.

@@

Another approach to shared control is allocating time slices or providing assistance
on an as-needed basis~\cite{IEEE07140030}.

This recent paper discusses shared control \url{http://www.ri.cmu.edu/pub_files/2015/7/Javdani15Hindsight.pdf}

@@

Point to DRC papers by multiple groups.

Figure~\ref{f:cga1} shows the architecture of our robot control based on hierarchical optimization. 
It was successfully tested on our SARCOS Primus Humanoid
and ATLAS (Boston Dynamics) platforms. Our ATLAS robot demonstrated rough terrain walking, ladder climbing, and full body manipulation in the DARPA Robotics Challenge (DRC) Trials of 2013. This work also received 
a ``Best Oral Paper Award'' at Humanoids 2013~\cite{Feng/etal13}. Figure~\ref{fig:terrain_pic} shows a sequence of snapshots of ATLAS 
walking on rough terrain and plots of measured trajectories of the feet and the center of mass (CoM) for the DRC Trials.
Figure~\ref{fig:ladder_pic} shows similar data for ladder climbing.
In the DARPA Robotics Challenge Finals in 2015 our robot was the only
biped that did not fall down or need to be rescued by 
humans during the two days of the evaluations~\cite{cga-hum15}.
Our robot performance was also the most consistent across the two days.
Because of the way the evaluations were run, we have only limited data
from the actual evaluations. Videos of our robot's performance during the
DRC Finals are available~\cite{cga-web-page}.

@@

The {\bf Multistep Low-Order Dynamics Planner} uses
receding horizon control to continually optimize trajectories of
motion and force of a simplified model of the current task.
In walking, this is usually center of mass ($COM_d(t)$),
angular momentum ($\mL_d(t)$),
contact force ($\mbox{contacts}_d(t)$) trajectories,
and value, heuristic cost, or utility functions ($V_c(t)$)
specifying tradeoffs between these quantities.
We optimize with a time horizon of several seconds
using Differential Dynamic Programming~\cite{ddp}.

For walking, our planner is similar in spirit to Kajita's Preview Control for a Linear Inverted
Pendulum Model (LIPM)~\cite{kajita03}. We are also using a CoM
model, reason about Zero Moment Point (ZMP), and use future information 
to guide the current trajectory. However, our planner can be generalized to 
nonlinear models and adjust foot steps while optimizing the CoM 
trajectory. 
%We explicitly add the vertical height in our  CoM model to handle height variations on rough terrain. 
Like capture point methods~\cite{capture_pt}, we take the 
next few steps into consideration but
do not plan to come to rest at the end. Ogura et al.~\cite{wabian2} 
investigated generating human-like walking with heel-strike and toe-off by 
parameterizing the swing foot trajectory. In contrast, we use very simple rules to guide the low level controller to achieve the same behaviors.

@@

For the QP, we use a formulation developed in our group 
\cite{Stephens-thesis,Whitman10,Whitman-thesis}. Unlike~\cite{mistry10,usc_id1} who 
use orthogonal decomposition to project the allowable motions into the null 
space of the constraint Jacobian, and minimize costs in the contact constraints and the commands, we directly 
optimize a quadratic cost in terms of state accelerations, torques and contact 
forces on the full robot model. This design choice allows us to trade off physical quantities of 
interest. We are also able to directly reason about inequality constraints 
such as center of pressure within the support polygon, friction, and torque 
limits. Although it becomes a bigger QP problem, we are still able to solve it 
in real time. Hutter el al.~\cite{eth_id} resolved redundancy in inverse 
dynamics using a kinematic task prioritization approach that ensures lower 
priority tasks always exist in the null space of higher priority ones. In 
contrast to their strictly hierarchical approach, we minimize a sum of 
weighted terms. 
We can directly specify the relative importance of the terms by adjusting 
the weights.

@@

Figure~\ref{f:cga1} does not include our state estimator, which provides
appropriate state information to all other modules. The details of this work on state estimation are
described in~\cite{xinji_com_est,Xinjilefu-thesis}.

@@

\begin{equation}
\dot{\vq}_d = \mJ^{-1} (\mK_1 \vf)
\label{eq:gootwcps}
\end{equation}
where $\dot{\vq}_d$ are the commanded exoskeleton joint velocities,
$\mJ$ is an appropriate Jacobian matrix, $\mK_1$ is a gain matrix,
and $\vf$ is the force vector to be controlled~\cite{IEEE06990981}.

@@

We have not seen a lot of discussion of state estimation of exoskeletons in
the literature.
In lieu of that,
we will discuss our approach to state estimation on humanoid robots here.
State estimation to deal with sensor noise,
disturbance estimation, and incomplete or
redundant sensing is a separate process and is as independent as possible
from the control method~\cite{certainty-eq,separation-prin}.
This makes evaluating different control methods much easier.
Disturbance observers for disturbances such as
moving support surfaces, contact or wind forces, and model errors
are also implemented.

@@

Joint velocities are estimated using actuator position sensors, actuator load
sensors, actuator commands, a single high
quality IMU at the pelvis, and multiple low quality (but cheap) MEMs IMUs
spread throughout the robot or exoskeleton. We also take into account feet and hand
contact forces, which
are measured. We use the full internal dynamics as the process model,
assuming the root dynamics are known. See Chapter 3 of~\cite{Xinjelefu-thesis}.
The internal dynamics are linearized and covariances updated
by a separate process at a 100Hz rate.
The joint positions are predicted by integrating the estimated velocities.

For higher level (overall) state estimation,
we developed a center of mass location and velocity estimator~\cite{}.
The center of mass filter is modeled as a multiple model EKF with contact switching
to handle changes in support: single support, double support, and off the ground.
We use the robot or exoskeleton
joint sensing (kinematics) to create a virtual sensor that measures
center of mass position and velocity relative to a stance foot.
We use Linear
Inverted Pendulum Model (LIPM) dynamics with an offset to predict
the center of mass motion. This offset can be interpreted
as a modelling error on the center of mass position, or an
external force exerted on the center of mass of the robot/exoskeleton,
or a combination of both. The center of mass estimator was
implemented on our Atlas humanoid robot. 
This estimator is especially helpful when compensating
for unplanned slow changing external forces applied at
unknown locations on the robot or exoskeleton, which is quite likely when
operating in tight spaces. It also handles relatively small
dynamic forces well when walking, e.g. dragging a tether
or pushing through a spring loaded door. Thanks to
the estimator, very little tuning is done for our mass model.
During the DARPA
Robotics Challenge Finals
where no safety belay was
allowed, our fall early warning system based on this estimator
successfully saved our robot from
falling on two occasions, and made us the only competitive team
without a fall or need for a human rescue among all teams.

{\bf Safety Code: Fall Prediction:}
The most significant contribution of the external force estimator is that it
can detect when a large external force is being applied that might
push the robot or exoskeleton over.
We compute a ``corrected capture point'' (CCP)~\cite{capture},
which is an offset to the current capture point.
The offset 
takes into account the estimated external force, represented as an offset
to the center of mass.
The corrected capture point getting close to the boundary of the polygon of
support warns the controller that the robot or exoskeleton might fall if the external
force is maintained.
We can also compute the corrected capture point assuming that the
external force follows a known time course plus an estimated constant offset,
or steadily increases or decreases for a fixed time interval
based on an estimated derivative.
We assume the external force is due to the robot's or exoskeleton's
movements, and not
due to external disturbances such as wind, a moving support platform
or external agents pushing on the robot.
The current
behavior is stopped and the robot or exoskeleton can safely be ``frozen'' in place.

A derivation of the corrected capture point starts with LIPM dynamics
augmented with a true center of mass offset and a true external force:
\begin{equation}
\ddot{c} = (c + c_{offset} + f_{ext}*z/mg - cop)*g/z
= (c + \Delta - cop)*g/z
\end{equation}
where $c$ is the location of the center of mass projected on the ground
plane,
$cop$ is the center of pressure
location in that ground plane, and $\Delta$ is the sum of the true center of mass
offset from the modeled center of mass and any external horizontal force.
Our extended Kalman filter estimates $\hat{c}$, $\hat{\dot{c}}$, and
$\hat{\Delta}$,
taking into account the
current center of mass height $z$. We assume a constant center of mass height
in estimating the corrected capture point based on the estimated capture point
as described in~\cite{capture}:
\begin{equation}
\widehat{CCP} = \widehat{CP} + \hat{\Delta} 
= \hat{c} + \hat{\dot{c}}\hat{z}/g + \hat{\Delta}
\end{equation}

@@

We see that Sensitivity Amplification Control uses acceleration as an input.
Estimating acceleration using double differentiation with low pass filtering
currently works well on real exoskeletons.
However, impacts, shock waves, and the fact that neither
the operator's body parts or the exoskeletons parts are rigid bodies and the
joints are not well defined suggest this approach can be improved with actual
acceleration measurements.
It is now possible to distribute MEMs IMUs throughout the exoskeleton to
improve both velocity and 
acceleration estimation~\cite{Xinjilefu-thesis}, which should improve
Sensitivity Amplification Control.

Disturbance observers~\cite{IEEE06197032}.

@@

This question needs to be addressed using simulation, and is deferred to a 
possible future white paper [robust.pdf].

@@

On the other hand, feedback control can lead to steady state or growing
oscillation, which we colloquially refer to as being unstable or instability.
Since walking and running are also oscillations which we refer to as stable
behaviors, the technical meaning of the word ``stable'' is more complex than
just ``it keeps moving'' or ``it doesn't do what the operator wants''.
In any event, we don't want oscillations.

@@

If the robustness of online optimization is not adequate, an alternative is
combining feedforward and feedback control into one function and using gradient
descent in simulation to improve the control robustness.
One technique to do this is multiple-model design~\cite{acc12}.

@@

{\it [Exoskeletons] have the implicit property
of causing a virtual modification of the dynamic response of
the human limb. We use this property of the exoskeletons
action to formulate a unified control design framework called
Integral Admittance [torque to angle] Shaping, which designs exoskeleton con-
trollers capable of producing the desired dynamic response
for the assisted limb. In this framework, a virtual increase
in the admittance of the limb is produced by coupling it
to an exoskeleton that exhibits active behavior. Specifically,
our framework shapes the magnitude profile of the integral
admittance (i.e. torque-to-angle relationship) of the coupled
human-exoskeleton system, such that the desired assistance is
achieved. This framework also ensures that the coupled stability
and passivity are guaranteed.}~\cite{Nagarajan_etal_2015}

{\it ... the impedance of the coupled human-exoskeleton
system needs to be reduced below that of the unassisted
human limb. This implies that the exoskeleton needs to
cancel its own impedance first and then compensate for
at least a part of the human limb’s impedance. Therefore,
the desired exoskeleton behavior must be that of a negative
impedance ... Consequently,
the feedback gains will all be
positive. In other words, the exoskeleton controller uses
positive feedback, and hence the exoskeleton exhibits active
behavior, which is capable of performing net positive work
on the limb. ...
However, positive feedback naturally raises the question of
stability, and so we now explain how coupled stability can be
achieved. Although the exoskeleton exhibits active behavior,
which can be potentially destabilizing, the controller can be
designed such that the coupled human-exoskeleton system
is stable and passive. ... 
}~\cite{Nagarajan_etal_2015}

{\it
... Inertia compensation is more complex ...
It can be shown that using only positive acceleration
feedback ..., the gain margin of the coupled system
reduces to the moment of inertia of the exoskeleton, which
implies that the exoskeleton controller ... can at the most
compensate for the exoskeleton’s own moment of inertia
before going unstable. This implies that the moment of
inertia of the coupled human-exoskeleton system cannot be
reduced below that of the unassisted human limb, without
compromising coupled stability. However, using low-pass
filtered acceleration feedback, it can be shown that inertia 
reduction can be achieved. ...
This work uses filtered acceleration
feedback with a second-order low-pass Butterworth filter. ...
}~\cite{Nagarajan_etal_2015}

>>>>

\begin{enumerate}
\item
We recommend using {\bf online optimization.} This architecture allows
a wide variety of control approaches to be implemented and work together.
\item
We also recommend a {\bf red team} that explores traditional 
{\bf frequency domain} control system
design.
\item
We recommend applying as {\bf many sensors} as possible, and asking the performers
to make it easy to add more by making the sensor network available in the design.
This maximizes the probability of success
with respect to control architectures and algorithms by enabling multiple control
approaches to be implemented, refined, and support each other.
\item
We recommend a {\bf ``symbiotic'' control system design} should be used as a backup,
in case more aggressive design philosophies such as ``invisible'' and ``natural''
control system designs are not achievable.
The ``symbiotic'' control system design
expects the operator to adapt to the exoskeleton and its control,
and has the control customized for the operator.
\end{enumerate}

@@

%\section{Symbiosis and Autonomy, Handling Errors, and Superhuman Reflexes}
A future white paper could discuss
possible combinations of response to user and exoskeleton
autonomy including balance control; response to trips, slips,
stumbles, and fumbles; task assistance and guidance (guide operator to
doorknob, button, or light switch); superhuman response to external
perturbations such as projectiles and explosions; and autonomous
execution (operator/exoskeleton symbiosis with multi-tasking)?

A future white paper could discuss how to handle exoskeleton
malfunctions and damage to the
system. Basically this boils down to fault detection, and switch over to
an appropriate safe mode of operation.

A future white paper could evaluate using simulation to assess
the various control approaches' robustness to modeling error and unmodeled
dynamics.

A future white paper could survey human learning with respect to exoskeletons
and teleoperation.\\
\url{https://www.newscientist.com/article/dn28155-exoskeleton-reveals-the-lengths-we-go-to-save-energy-when-walking/}\\

\begin{verbatim}
K. E. Gordon, C. R. Kinnaird, and D. P. Ferris, “Locomotor adaptation
to a soleus emg-controlled antagonistic exoskeleton,” J. Neurophysiol.,
vol. 109, no. 7, pp. 1804–1814, 2013.
\end{verbatim}

A future white paper could survey adaptive control of exoskeletons.\\
{\it
The relationship between physical human-exoskeleton interaction and
dynamic factors: using a learning approach for control applications}\\
\url{http://info.scichina.com/sciFe/EN/abstract/abstract516165.shtml}

A future white paper could survey the role of phase estimation, behavioral clocks,  phase reset events such as impacts, central pattern generators,
and oscillator approaches in robot and exoskeleton control.

\subsection{Related Surveys}

Active Exoskeleton Control Systems: State of the Art (2012)\\
\url{http://www.sciencedirect.com/science/article/pii/S1877705812026732}\\
Our copy:\\
\url{http://www.cs.cmu.edu/~cga/exo/info/Anam2012.pdf}

@@

If you ask people who research complex computationally expensive
high end control
systems for a living how to control a system,
they will recommend complex computationally expensive high end control systems.
It is entirely possible that simple local low dimensional controllers
designed using traditional (frequency domain) techniques will do
quite well, especially if the appropriate sensors (such as 
operator-exoskeleton force sensors) are available.
It may even be the case that frequency domain techniques will handle
structural modeling error (jello-like human operators, deformable materials
(straps and pads, wires and hoses), and structural deformation and joint
play in lightweight frames) better than the state space oriented 
and other advanced techniques advocated here.

We recommend a red team that explores traditional 
frequency domain control system
design, as well as an emphasis on advanced control approaches.

@@

Typically low level control is implemented
by a high sampling rate global servo [Boston Dynamics Atlas humanoid]
or distributed limb or joint-level high sampling rate servos
[Sarcos Primus humanoid].
Often, this controller is provided by the manufacturer of the robot
or exoskeleton and runs on a separate computer system from the
``user's'' or task/application controller.

@@

In this architecture, state estimation to deal with sensor noise,
disturbance estimation, and incomplete or
redundant sensing is a separate process and is as independent as possible
from the control method.
%~\cite{certainty-eq,separation-prin}.
This makes evaluating different control methods much easier.
Disturbance observers for disturbances such as
moving support surfaces, contact or wind forces, and model errors
are also implemented.

@@

The recommended control architecture is intuitive, abstracts the hardware 
and lower level control in a straightforward way, is easy to program, and is
commonly used. Variants of this architecture were used by almost all teams
in the DARPA Robotics Challenge.
%~\cite{}.

@@

An alternative approach to state estimation is to design state estimation
and control laws simultaneously.
%~\cite{LQG-LTR}. 
Another alternative is
to have a controller with internal state, and design or learn a policy
that works well, but does not have an explicit state estimation 
part.
%~\cite{lead-lag}.

@@

A key question is where does the desired acceleration $\ddot{\vq}_d$ come from?
This is discussed in a companion paper.
\url{http://www.cs.cmu.edu/~cga/exo/intent.pdf}

@@

Full (operator) body
{\bf online inverse dynamics} as described in Section~\ref{sec:invdyn}
no longer works if $\mJ_{\myx}$ is not invertible (not square or not full rank).
In the underactuated and overactuated cases $\mJ_{\myx}$ is not square. 
For a partial (operator) body
exoskeleton we can rewrite the dynamics in terms of only those
joints the exoskeleton actuates, and invert those dynamics to cancel exoskeleton
effects on those joints. This is what has been done with the BLEEX exoskeleton,
for example.
%~\cite{}.

@@

Another approach to solving this problem is prioritized control, which does
not include optimization.
%~\cite{}
For reasons we will describe, we believe
online optimization is much more flexible and can handle programming exoskeleton
behavior much better, so we will not discuss prioritized control further.

@@

Quadratic programming
%~\cite{Wikipedia} 
solves the following problem:
Minimize
\begin{equation}
\vx^{\tr} \mQ \vx + c^{\tr} \vx
\end{equation}
with inequality constraints
\begin{equation}
\mA_1 \vx \le \vb_1
\end{equation}
and equality constraints
\begin{equation}
\mA_2 \vx = \vb_2
\end{equation}
The optimization variable $\vx$ includes accelerations $\ddot{\vq}$, actuator
commands $\vu$, and contact forces $\vf$.
We could optimize only taking into account the exoskeleton and its dynamics,
minimizing the discrepancy between actual and desired accelerations, actuator command
magnitude, and contact force magnitude:
\begin{equation}
C_{\myx}(\vx) = (\vq - \vq_d)^{\tr} \mW_{\vq} (\vq - \vq_d)
+ \vu_{\myx}^{\tr} \mW_{\vu_{\myx}} \vu_{\myx}
+ \vf_{\ox}^{\tr} \mW_{\vf_{\ox}} \vf_{\ox}
+ \vf_{\xw}^{\tr} \mW_{\vf_{\xw}} \vf_{\xw}
\end{equation}
$\mW_{\vq}$, $\mW_{\vu_{\myx}}$, $\mW_{\vf_{\ox}}$,
and $\mW_{\vf_{\xw}}$ are weight matrices to trade off the
different errors/magnitudes and error/magnitude directions.
The dynamics (equation~\ref{eq:xdynamics2}) are enforced as an equality constraint.
Actuator limits, friction limits, center of pressure location constraints,
and unidirectional contact constraints are all
enforced as inequality constraints.
Variants of this optimization
were performed every 1-2ms by various teams to control the
Atlas robots in the DARPA Robotics Challenge.

@@

Another approach to shared control is allocating time slices or providing assistance
on an as-needed basis.\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=7140030}

This recent paper discusses shared control\\
\url{http://www.ri.cmu.edu/pub_files/2015/7/Javdani15Hindsight.pdf}

@@

More details on our work and pointers to other work are in the papers at:\\
\url{http://www.cs.cmu.edu/~cga/drc}.

Figure~\ref{f:cga1} shows the architecture of our robot control based on hierarchical optimization. 
It was successfully tested on our SARCOS Primus Humanoid
and ATLAS (Boston Dynamics) platforms. Our ATLAS robot demonstrated rough terrain walking, ladder climbing, and full body manipulation in the DARPA Robotics Challenge (DRC) Trials of 2013. This work also received 
a ``Best Oral Paper Award'' at Humanoids 2013. Figure~\ref{fig:terrain_pic} shows a sequence of snapshots of ATLAS 
walking on rough terrain and plots of measured trajectories of the feet and the center of mass (CoM) for the DRC Trials.
Figure~\ref{fig:ladder_pic} shows similar data for ladder climbing.
In the DARPA Robotics Challenge Finals in 2015 our robot was the only
biped that did not fall down or need to be rescued by 
humans during the two days of the evaluations.
Our robot performance was also the most consistent across the two days.
Because of the way the evaluations were run, we have only limited data
from the actual evaluations. Videos of our robot's performance during the
DRC Finals are available.\\
\url{http://www.cs.cmu.edu/~cga/drc}.

@@

The {\bf Multistep Low-Order Dynamics Planner} uses
receding horizon control to continually optimize trajectories of
motion and force of a simplified model of the current task.
In walking, this is usually center of mass ($COM_d(t)$),
angular momentum ($\mL_d(t)$),
contact force ($\mbox{contacts}_d(t)$) trajectories,
and value, heuristic cost, or utility functions ($V_c(t)$)
specifying tradeoffs between these quantities.
We optimize with a time horizon of several seconds
using Differential Dynamic Programming.
%~\cite{ddp}.

For walking, our planner is similar in spirit to Kajita's Preview Control for a Linear Inverted
Pendulum Model (LIPM)
%~\cite{kajita03}.
We are also using a CoM
model, reason about Zero Moment Point (ZMP), and use future information 
to guide the current trajectory. However, our planner can be generalized to 
nonlinear models and adjust foot steps while optimizing the CoM 
trajectory. 
%We explicitly add the vertical height in our  CoM model to handle height variations on rough terrain. 
Like capture point methods,
%~\cite{capture_pt}, 
we take the 
next few steps into consideration but
do not plan to come to rest at the end. Ogura et al.
%~\cite{wabian2} 
investigated generating human-like walking with heel-strike and toe-off by 
parameterizing the swing foot trajectory. In contrast, we use very simple rules to guide the low level controller to achieve the same behaviors.

@@

For the QP, we use a formulation developed in our group.
%\cite{Stephens-thesis,Whitman10,Whitman-thesis}. 
Unlike those
% ~\cite{mistry10,usc_id1}
who use orthogonal decomposition to project the allowable motions into the null 
space of the constraint Jacobian, and minimize costs in the contact constraints and the commands, we directly 
optimize a quadratic cost in terms of state accelerations, torques and contact 
forces on the full robot model. This design choice allows us to trade off physical quantities of 
interest. We are also able to directly reason about inequality constraints 
such as center of pressure within the support polygon, friction, and torque 
limits. Although it becomes a bigger QP problem, we are still able to solve it 
in real time. Hutter el al.
%~\cite{eth_id}
resolved redundancy in inverse 
dynamics using a kinematic task prioritization approach that ensures lower 
priority tasks always exist in the null space of higher priority ones. In 
contrast to their strictly hierarchical approach, we minimize a sum of 
weighted terms. 
We can directly specify the relative importance of the terms by adjusting 
the weights.

@@

Figure~\ref{f:cga1} does not include our state estimator, which provides
appropriate state information to all other modules. The details of this work on state estimation are
described in.
\url{http://www.cs.cmu.edu/~cga/drc}.
%~\cite{xinji_com_est,Xinjilefu-thesis}.

@@

\begin{equation}
\dot{\vq}_d = \mJ^{-1} (\mK_1 \vf)
\label{eq:gootwcps}
\end{equation}
where $\dot{\vq}_d$ are the commanded exoskeleton joint velocities,
$\mJ$ is an appropriate Jacobian matrix, $\mK_1$ is a gain matrix,
and $\vf$ is the force vector to be controlled.\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=6990981}
%~\cite{IEEE06990981}.

@@

We have not seen a lot of discussion of state estimation of exoskeletons in
the literature.
In lieu of that,
we will discuss our approach to state estimation on humanoid robots here.
State estimation to deal with sensor noise,
disturbance estimation, and incomplete or
redundant sensing is a separate process and is as independent as possible
from the control method
%~\cite{certainty-eq,separation-prin}.
This makes evaluating different control methods much easier.
Disturbance observers for disturbances such as
moving support surfaces, contact or wind forces, and model errors
are also implemented.

@@

Joint velocities are estimated using actuator position sensors, actuator load
sensors, actuator commands, a single high
quality IMU at the pelvis, and multiple low quality (but cheap) MEMs IMUs
spread throughout the robot or exoskeleton. We also take into account feet and hand
contact forces, which
are measured. We use the full internal dynamics as the process model,
assuming the root dynamics are known. See Chapter 3 of\\
\url{http://www.cs.cmu.edu/~xxinjile/pdf/main.pdf}\\
%~\cite{Xinjelefu-thesis}.
The internal dynamics are linearized and covariances updated
by a separate process at a 100Hz rate.
The joint positions are predicted by integrating the estimated velocities.

For higher level (overall) state estimation,
we developed a center of mass location and velocity estimator.
%~\cite{}.
The center of mass filter is modeled as a multiple model EKF with contact switching
to handle changes in support: single support, double support, and off the ground.
We use the robot or exoskeleton
joint sensing (kinematics) to create a virtual sensor that measures
center of mass position and velocity relative to a stance foot.
We use Linear
Inverted Pendulum Model (LIPM) dynamics with an offset to predict
the center of mass motion. This offset can be interpreted
as a modelling error on the center of mass position, or an
external force exerted on the center of mass of the robot/exoskeleton,
or a combination of both. The center of mass estimator was
implemented on our Atlas humanoid robot. 
This estimator is especially helpful when compensating
for unplanned slow changing external forces applied at
unknown locations on the robot or exoskeleton, which is quite likely when
operating in tight spaces. It also handles relatively small
dynamic forces well when walking, e.g. dragging a tether
or pushing through a spring loaded door. Thanks to
the estimator, very little tuning is done for our mass model.
During the DARPA
Robotics Challenge Finals
where no safety belay was
allowed, our fall early warning system based on this estimator
successfully saved our robot from
falling on two occasions, and made us the only competitive team
without a fall or need for a human rescue among all teams.

{\bf Safety Code: Fall Prediction:}
The most significant contribution of the external force estimator is that it
can detect when a large external force is being applied that might
push the robot or exoskeleton over.
We compute a ``corrected capture point'' (CCP),
%~\cite{capture},
which is an offset to the current capture point.
The offset 
takes into account the estimated external force, represented as an offset
to the center of mass.
The corrected capture point getting close to the boundary of the polygon of
support warns the controller that the robot or exoskeleton might fall if the external
force is maintained.
We can also compute the corrected capture point assuming that the
external force follows a known time course plus an estimated constant offset,
or steadily increases or decreases for a fixed time interval
based on an estimated derivative.
We assume the external force is due to the robot's or exoskeleton's
movements, and not
due to external disturbances such as wind, a moving support platform
or external agents pushing on the robot.
The current
behavior is stopped and the robot or exoskeleton can safely be ``frozen'' in place.

A derivation of the corrected capture point starts with LIPM dynamics
augmented with a true center of mass offset and a true external force:
\begin{equation}
\ddot{c} = (c + c_{offset} + f_{ext}*z/mg - cop)*g/z
= (c + \Delta - cop)*g/z
\end{equation}
where $c$ is the location of the center of mass projected on the ground
plane,
$cop$ is the center of pressure
location in that ground plane, and $\Delta$ is the sum of the true center of mass
offset from the modeled center of mass and any external horizontal force.
Our extended Kalman filter estimates $\hat{c}$, $\hat{\dot{c}}$, and
$\hat{\Delta}$,
taking into account the
current center of mass height $z$. We assume a constant center of mass height
in estimating the corrected capture point based on the estimated capture point:
% as described in~\cite{capture}:
\begin{equation}
\widehat{CCP} = \widehat{CP} + \hat{\Delta} 
= \hat{c} + \hat{\dot{c}}\hat{z}/g + \hat{\Delta}
\end{equation}

@@

We see that Sensitivity Amplification Control uses acceleration as an input.
Estimating acceleration using double differentiation with low pass filtering
currently works well on real exoskeletons.
However, impacts, shock waves, and the fact that neither
the operator's body parts or the exoskeletons parts are rigid bodies and the
joints are not well defined suggest this approach can be improved with actual
acceleration measurements.
It is now possible to distribute MEMs IMUs throughout the exoskeleton to
improve both velocity and 
acceleration estimation
%~\cite{Xinjilefu-thesis}, 
which should improve
Sensitivity Amplification Control.

% Disturbance observers~\cite{IEEE06197032}.

@@

This question needs to be addressed using simulation, and is deferred to a 
possible future white paper.

@@

On the other hand, feedback control can lead to steady state or growing
oscillations, which we colloquially refer to as being unstable or an
instability.
Since walking and running are also oscillations which we refer to as stable
behaviors, the technical meaning of the word ``stable'' is more complex than
just ``it keeps moving'' or ``it doesn't do what the operator wants''.
In any event, we don't want oscillations.

@@

If the robustness of online optimization is not adequate, an alternative is
combining feedforward and feedback control into one function and using gradient
descent in simulation to improve the control robustness.
One technique to do this is multiple-model design.
%~\cite{acc12}.
\url{http://www.cs.cmu.edu/~cga/papers/acc2012-tr.pdf}

@@

Integral admittance control is a refinement of 
sensitivity amplification control.
This section is quotes from:\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=7139989}

{\it [Exoskeletons] have the implicit property
of causing a virtual modification of the dynamic response of
the human limb. We use this property of the exoskeletons
action to formulate a unified control design framework called
Integral Admittance [torque to angle] Shaping, which designs exoskeleton con-
trollers capable of producing the desired dynamic response
for the assisted limb. In this framework, a virtual increase
in the admittance of the limb is produced by coupling it
to an exoskeleton that exhibits active behavior. Specifically,
our framework shapes the magnitude profile of the integral
admittance (i.e. torque-to-angle relationship) of the coupled
human-exoskeleton system, such that the desired assistance is
achieved. This framework also ensures that the coupled stability
and passivity are guaranteed.}
%~\cite{Nagarajan_etal_2015}

{\it ... the impedance of the coupled human-exoskeleton
system needs to be reduced below that of the unassisted
human limb. This implies that the exoskeleton needs to
cancel its own impedance first and then compensate for
at least a part of the human limb’s impedance. Therefore,
the desired exoskeleton behavior must be that of a negative
impedance ... Consequently,
the feedback gains will all be
positive. In other words, the exoskeleton controller uses
positive feedback, and hence the exoskeleton exhibits active
behavior, which is capable of performing net positive work
on the limb. ...
However, positive feedback naturally raises the question of
stability, and so we now explain how coupled stability can be
achieved. Although the exoskeleton exhibits active behavior,
which can be potentially destabilizing, the controller can be
designed such that the coupled human-exoskeleton system
is stable and passive. ... 
}
%~\cite{Nagarajan_etal_2015}
